---
title: "Exploratory Data Analysis"
author: "Kamil Szymkowski"
date: "15/01/2021"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data file

```{r file, echo = FALSE}
ships <- read.csv("./ships.csv")
class(ships)
utils:::format.object_size(file.info("./ships.csv")$size, "auto")
dim(ships)
```

The `ships.csv` is an example of a large dataset. The file takes up 129 MB on the disk and consists of over 1 million cases. Each case is described using 20 variables.


## Variables
```{r head, echo = FALSE}
head(ships)
```

The variables can be grouped into the following categries:

* Static information:
  + SHIP_ID - ship’s unique identifier
  + SHIPNAME - ship’s name
  + SHIPTYPE - ship’s type
  + Ship_type - ship’s type from SHIPTYPE
  + WIDTH - ship’s width in meters
  + LENGTH - ship’s length in meters
  + FLAG - ship’s flag
  + DWT - ship’s deadweight in tones
* Dynamic information:
  + LAT - ship’s latitude
  + LON - ship’s longitude
  + DATETIME - date and time of the observation
  + Date - date extracted from DATETIME
  + Week_nb - week number extracted from date
  + COURSE - ship’s course as angle
  + SPEED - ship’s speed in knots
  + HEADING - ship’s compass direction
  + Is_parked - indicator whether the ship is moving or not 
  + DESTINATION - ship’s destination (reported by the crew)
  + PORT - current port reported by the vessel
  + Port - current port assigned based on the ship’s location


```{r variables}
str(ships)
summary(ships)
```

